---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
# bnclassify

[![Travis-CI Build Status](https://travis-ci.org/bmihaljevic/bnclassify.svg?branch=master)](https://travis-ci.org/bmihaljevic/bnclassify)
[![codecov.io](https://codecov.io/github/bmihaljevic/bnclassify/coverage.svg?branch=master)](https://codecov.io/github/bmihaljevic/bnclassify?branch=master)

The bnclassify package implements algorithms for learning discrete Bayesian network classifiers from data and estimating their predictive accuracy.

# Example 

Let us load a data set and learn a tree-augmented naive Bayes by maximizing the BIC score.
```{r}
library(bnclassify)
data(car)
tn <- tan_cl('class', car, score = 'bic')
tn
plot(tn)
```

After we fit its parameters, we can use it to predict class labels. 
```{r}
tn <- lp(tn, car, smooth = 0.01)
p <- predict(tn, car, prob = TRUE)
head(p)
p <- predict(tn, car, prob = FALSE)
head(p)
```

Estimate predictive accuracy with cross validation.
```{r}
cv(tn, car, k = 10, smooth = 0.01, dag = TRUE)
```

# Install

Make sure you have at least version 3.2.0 of R. You will need to install packages from Bioconductor.  

```{r, eval = FALSE}
source("http://bioconductor.org/biocLite.R")
biocLite(c("graph", "RBGL", "Rgraphviz"))
```

When you read this bnclassify may already be on CRAN. In that case, use:
```{r, eval = FALSE}
install.packages('bnclassify')
```

If not, you can install the current version from github:
```{r, eval = FALSE}
install.packages('devtools')
devtools::install_github('bmihaljevic/bnclassify', build_vignettes = TRUE)
#devtools::install_github('bmihaljevic/bnclassify')
```

# Overview

See the list of implemented algorithms with 

```{r, eval = FALSE}
?bnclassify
```

Use the vignette to get started. 
```{r, eval = FALSE}
vignette('bnclassify')
```
