---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```
# bnclassify

[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/bnclassify)](https://cran.r-project.org/package=bnclassify)
![](https://cranlogs.r-pkg.org/badges/bnclassify?color=yellow)
![](https://cranlogs.r-pkg.org/badges/grand-total/bnclassify?color=yellowgreen)
[![Research software impact](http://depsy.org/api/package/cran/bnclassify/badge.svg)](http://depsy.org/package/r/bnclassify)
[![Codecov test coverage](https://codecov.io/gh/bmihaljevic/bnclassify/branch/master/graph/badge.svg)](https://app.codecov.io/gh/bmihaljevic/bnclassify?branch=master)
[![R-CMD-check](https://github.com/bmihaljevic/bnclassify/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bmihaljevic/bnclassify/actions/workflows/R-CMD-check.yaml)

Implements algorithms for learning discrete Bayesian network classifiers from data, as well as functions for using these classifiers for prediction, assessing their predictive performance, and inspecting and analyzing their properties.

# Example 

Load a data set and learn a one-dependence estimator by maximizing Akaike's information criterion (AIC) score.
```{r}
library(bnclassify)
data(car)
tn <- tan_cl('class', car, score = 'aic')
tn
plot(tn)
```

After learning the network's parameters, you can use it to classify data. 
```{r}
tn <- lp(tn, car, smooth = 0.01)
p <- predict(tn, car, prob = TRUE)
head(p)
p <- predict(tn, car, prob = FALSE)
head(p)
```

Estimate predictive accuracy with cross validation.
```{r}
cv(tn, car, k = 10)
```

Or compute the log-likelihood 
```{r}
logLik(tn, car)
```

# Install

Make sure you have at least version 3.2.0 of R. You can install `bnclassify` from CRAN:
```{r, eval = FALSE}
install.packages('bnclassify')
```

Or get the current development version from Github:
```{r, eval = FALSE}
# install.packages('devtools')
devtools::install_github('bmihaljevic/bnclassify')
# devtools::install_github('bmihaljevic/bnclassify', build_vignettes = TRUE)
```

Ideally, you would use the `build_vignettes = TRUE` version, and thus get the vignettes, but it requires programs such as texi2dvi to be installed on your side. 

# Overview

See an overview of the package and examples of usage:

```{r} 
vignette('overview', package = 'bnclassify')
```


See the list of main functionalities.

```{r, eval = FALSE}
?bnclassify
```

Use the usage vignette for more details on the functions.

``` r
vignette('usage', package = 'bnclassify')
```

Then have a look at the remaining vignettes.

``` r
browseVignettes("bnclassify")
```
