---
title: "bnclassify"
author: "Bojan Mihaljevic, Concha Bielza, Pedro Larranaga"
date: "`r Sys.Date()`"
output: 
  rmarkdown::pdf_document:
    toc: true
fontsize: 11pt
vignette: >
  %\VignetteIndexEntry{bnclassify}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Predicting 

### 0 probabilities
If for some instance there is 0 probability for each class, then a uniform distribution over the classes is returned (not the class prior). 

```{r}
library(bnclassify)
data(car)
nb <- nb('class', car)
nb <- lp(nb, car[c(1, 700), ], smooth=0)
predict(object=nb, newdata=car[1000:1001, ], prob = TRUE)
```

### Speed 
It is much faster than gRain and identical to bnlearn.

```{r}
library(microbenchmark)
nb <- nb('class', car)
nb <- lp(nb, car, smooth = 0)
gr <- to_grain(nb)
microbenchmark(predict(object = nb, newdata = car, prob = TRUE))
microbenchmark(gRain::predict.grain(gr, 'class', newdata = car),
                               times = 1)
```

Note that when predicting on a data set with incomplete cases, gRain is used underneath and it will be slow


## Cross-validation
There is rather fast cross-validation. It takes advantage of the modularity of Bayesian networks by accessing only onces the CPTs shared accross structures. 

```{r}
a <- lp(nb('class', car), car, smooth = 1)	
b <- lp(nb('class', car[, 'class', drop = FALSE]), car, smooth = 1)
d <- lp(nb('class', car[, c(sample(1:6, 4), 7), drop = FALSE]), car, smooth = 1)	
set.seed(0)
microbenchmark(r <- dag_cv(list(a, b, d), 'class', car, smooth = 1, k = 10))
r
```

## Feature selection

Some algorithms perform implicit feature selection. E.g., ... For more, use the mlr or feature selector package. See below. 

## mlr package
It's easy to use bnclassify with the mlr package. If you have mlr installed, you just need to call `as_mlr()` to use mlr functions: select features, resample, etc.

```{r}
  library(mlr)
  ctrl = makeFeatSelControlSequential(alpha = 0, method = "sfs")
  rdesc = makeResampleDesc(method = "Holdout")
  ct <- mlr::makeClassifTask(id = "compare", data = car, target = 'class', 
                            fixup.data = "no", check.data = FALSE)  
  nf <- lp(nb('class', car), car, 1)
  bnl <- as_mlr(nf, dag = TRUE)
  sfeats = selectFeatures(learner = bnl, task = ct, resampling = rdesc,
                          control = ctrl, show.info = FALSE)
  sfeats$x
  detach('package:mlr')
```



## Vignette docs

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` setion of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
