---
title: "`bnclassify`: Learning Bayesian Network Classifiers"  
author: "Bojan Mihaljević, Concha Bielza and Pedro Larrañaga"
date: "`r Sys.Date()`"  
output:
  rmarkdown::pdf_document:
    number_sections: true
    keep_tex: true
    includes:
          in_header: header.tex
    citation_package: natbib 
bibliography: bnclassify.bib  
fontsize: 11pt
vignette: >
  %\VignetteIndexEntry{overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
--- 

```{r child = 'includes/abstract.rmd'} 
``` 
\tableofcontents   
```{r child = 'includes/intro.rmd'} 
``` 
The rest of this paper is structured as follows. Section \ref{sec:bcground} provides background on Bayesian network classifiers. Section \ref{sec:functionalities} describes the implemented functionalities. Section \ref{sec:usage} illustrates usage with a synthetic data set. Section \ref{sec:implementation} discusses implementation while Section \ref{sec:relatedsw} briefly surveys related software. Finally, Section \ref{sec:conclusion} concludes and outlines future work.   
```{r child = 'includes/background.rmd'}
``` 
```{r child = 'includes/functionalities.rmd'}
``` 
```{r child = 'includes/usage.rmd'}
``` 
```{r child = 'includes/impl-related-conclusion.rmd'}
```
